### 2016-01-18 paper.tex
Remove paragraph about solidity and clean up later motivation a bit

The performance of cross-flow MHK turbines thus depends on both Reynolds number
and solidity (note that these issues are related, since an average blade chord
Reynolds number, $Re_{c,\mathrm{avg}} \approx \lambda U_\infty c/ \nu$, can be
expressed in terms of tip speed ratio $\lambda$, the optimal value of which is a
function of solidity). If numerical models are validated with physical model
data that was obtained at insufficiently high Reynolds numbers, it cannot be
determined whether problems with model predictions are caused by Reynolds number
effects, issues related to higher solidity, or both. It is uncertain whether
numerical models validated with physical model data obtained at low Reynolds
number should be considered validated at all, since the scale at which the model
will be applied is orders of magnitude larger. One way to overcome this
uncertainty is to show that the scaled physical model test has become Reynolds
number independent, and therefore validation efforts should be relevant at
full-scale.

For example, the effect of Reynolds number on average power output was shown to
be significant on the 2 m Sandia Research Darrieus turbine in wind tunnel
testing \cite{Blackwell1976}: Maximum power coefficient, $C_{P_{\max}}$,
increased with Reynolds number, $Re_c$, along with a shift of the location of
$C_{P_{\max}}$ toward lower tip speed ratios due to delayed blade stall. The
effects of Reynolds number were quite dramatic over a relatively small range of
$Re_c \approx 1.1 \times 10^5$--$2.9 \times 10^5$. More recently, Bachant and
Wosnik \cite{Bachant2014, Bachant2015-RVAT-Re-dep} showed that performance and
near-wake characteristics of a high solidity cross-flow turbine become Reynolds
number independent at $Re_c \approx 2 \times 10^5$.

The need for experimental data that is relevant to full-scale behavior stems
from the need to validate numerical models---most importantly mid-fidelity
models---desirable for MHK developers to predict the performance of their
cross-flow turbine designs, since physical modeling at appropriate scales can be
prohibitively expensive in the early stages of engineering.

To date, attempts to validate SNL's mid-fidelity CACTUS vortex line model
\cite{Murray2011} have relied on measurements from the Saint Anthony Falls
Laboratory (SAFL) \cite{Hill2014} and the University of New Hampshire (UNH)
\cite{Neary2013, Michelen2014}. For the SAFL experiments (RM2), the chord
Reynolds number, $Re_c \sim 10^4$, was below the threshold needed to properly
simulate lift and stall characteristics. For the UNH experiment
\cite{Bachant2013}, the chord Reynolds number was sufficiently high at $Re_c
\approx 2.7 \times 10^5$, but the chord/radius ratio and solidity (13.4\%)
created instability in the free-wake evolution in the model, which caused
significant overestimation of power coefficient.

>>>>

The effect of Reynolds number on average power output was shown to be
significant on the 2 m Sandia Research Darrieus turbine in wind tunnel testing
\cite{Blackwell1976}: Maximum power coefficient, $C_{P_{\max}}$, increased with
Reynolds number, $Re_c$ (based on blade chord rather than diameter, though these
are approximately proportional), along with a shift of the location of
$C_{P_{\max}}$ toward lower tip speed ratios due to delayed blade stall. The
effects of Reynolds number were quite dramatic over a relatively small range of
$Re_c \approx 1.1 \times 10^5$--$2.9 \times 10^5$. More recently, Bachant and
Wosnik \cite{Bachant2014, Bachant2015-RVAT-Re-dep} showed that performance and
near-wake characteristics of a high solidity cross-flow turbine become Reynolds
number independent at $Re_c \approx 2 \times 10^5$.

As part of the engineering process it is generally less expensive to assess
designs via numerical rather than physical models. However, it is
important---especially when dealing with fluid dynamics where the ``exact''
physics cannot be resolved with even the most advanced computers---that
numerical models be validated against experimental data. It is uncertain whether
numerical models validated with physical model data obtained at low Reynolds
number should be considered validated at all, since the scale at which the model
will be applied for real world design problems is orders of magnitude larger.
One way to overcome this uncertainty is to show that the scaled physical model
test has become Reynolds number independent, and therefore validation efforts
should be relevant at full-scale, which was the strategy employed here.
